---
title: "student-teacher-ratios"
author: "DeFrayne"
date: "June 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
library(WDI)
theme_set(theme_light())
#Import the student to teacher ratio data
student_ratio <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-07/student_teacher_ratio.csv")

```

```{r}
#Exploring the dataset
View(student_ratio)

colnames(student_ratio)

student_ratio %>% count(flags, sort=T)

student_ratio %>% count(year, sort=T)

student_ratio %>% count(indicator, sort=T)

student_ratio_2015 <- student_ratio %>%
  filter(indicator == "Primary Education", year==2015, !is.na(student_ratio))

student_ratio_2015 %>%
  arrange(desc(student_ratio)) %>%
  slice(c(1:10, seq(n() - 10, n()))) %>%
  mutate(country=fct_reorder(country, student_ratio)) %>%
  ggplot(aes(country,student_ratio)) +
  geom_col() +
  coord_flip() +
  labs(title="Student to Teacher Ratio: Top 10 and Bottom 10 Countries",
       y="Student to Teacher Ratio",
       x="")

student_ratio_2015 %>%
  ggplot(aes(student_ratio)) +
  geom_histogram() +
  scale_x_log10()
```
There is an apparent negative correlation between national wealth and student/teacher ratio. There are some clear exceptions, such as Cuba having a low student/teacher ratio and the United States not being in the top 10.

```{r}
WDIsearch("gdp per capita") %>% View()

WDIsearch("literacy") %>% View()

WDIsearch("public.*education") %>% View()

WDIsearch("population") %>%
  as.data.frame() %>%
  tbl_df() %>%
  filter(str_detect(name,"^Population"))

indicators_raw <- WDI(indicator=c("NY.GDP.PCAP.CD",
                                  "SP.POP.TOTL",
                                  "SE.ADT.LITR.ZS",
                                  "SE.XPD.TOTL.GD.ZS",
                                  "SE.SEC.NENR.MA",
                                  "SE.SEC.NENR.FE"
                                  ),
                      start=2015,end=2015, extra=T) %>%
  tbl_df()

indicators <- indicators_raw %>%
  select(country_code=iso3c,
         region,
         NY.GDP.PCAP.CD:SE.SEC.NENR.FE) %>%
  mutate(country_code = as.character(country_code))


student_ratio_2015 %>%
  inner_join(indicators, by="country_code") %>%
  arrange(desc(SP.POP.TOTL)) %>%
  ggplot(aes(NY.GDP.PCAP.CD, student_ratio)) +
  geom_point(aes(size=SP.POP.TOTL, color=region)) +
  scale_x_log10() +
  scale_y_log10() + 
  scale_size_continuous(labels=scales::comma_format(), range=c(0.25,12)) +
  geom_text(aes(label=country), vjust=1, hjust=1, check_overlap=T) +
  labs(x="GDP per Capita",
       y="Student to Teacher Ratio in Primary Education",
       color="Region",
       title="Visualized negative correlation between GDP per capita and student/teacher ratio",
       size="Population")
```
This confirms that there exists a negative correlation between national wealth and student/teacher ratio.

```{r}
WDIsearch("education") %>%
  as.data.frame() %>%
  tbl_df() %>%
  filter(str_detect(name,"population"))

```
#Primary vs. Secondary Education
```{r}
secondary_primary_education <- student_ratio %>%
  filter(year==2015,
         !is.na(student_ratio),
         indicator %in% c("Primary Education", "Secondary Education")) %>%
  group_by(country) %>%
  filter(n()==2) %>%
  ungroup()

secondary_primary_education %>%
  inner_join(indicators, by="country_code") %>%
  arrange(desc(SP.POP.TOTL)) %>%
  ggplot(aes(NY.GDP.PCAP.CD, student_ratio)) +
  geom_point(aes(size=SP.POP.TOTL, color=region)) +
  geom_text(aes(label=country), vjust=1, hjust=1, check_overlap=TRUE) +
  scale_x_log10() +
  scale_y_log10() +
  scale_size_continuous(labels = scales::comma_format(), range=c(0.25,12)) +
  facet_wrap(~ indicator) +
  labs(x = "GDP per Capita",
       y = "Student to Teacher Ratio in Primary Education",
       subtitle = "In 2015",
       title = "GDP per Capita and Student/Teacher Ratio are Negatively Correlated",
       color = "Region",
       size = "Population")
```
#Other Indicators
```{r}
student_ratio_2015 %>%
  inner_join(indicators, by="country_code") %>%
  filter(!is.na(SE.ADT.LITR.ZS)) %>%
  mutate(literacy=SE.ADT.LITR.ZS/100) %>%
  ggplot(aes(NY.GDP.PCAP.CD,literacy)) +
  geom_point() +
  geom_text(aes(label=country), vjust=1, hjust=1, check_overlap=TRUE) +
  scale_x_log10() + 
  scale_y_continuous(labels=scales::percent_format())

student_ratio_2015 %>%
  inner_join(indicators, by="country_code") %>%
  filter(!is.na(SE.ADT.LITR.ZS)) %>%
  mutate(literacy=SE.ADT.LITR.ZS/100) %>%
  ggplot(aes(student_ratio,literacy)) +
  geom_point() +
  geom_text(aes(label=country), vjust=1, hjust=1, check_overlap=TRUE) +
  scale_x_log10() + 
  scale_y_continuous(labels=scales::percent_format())
```

#Exploring secondary education enrollment rates
```{r}
joined <- student_ratio_2015 %>%
  inner_join(indicators, by="country_code") %>%
  mutate(secondary_enrollment = (SE.SEC.NENR.FE+SE.SEC.NENR.MA)/2)

joined %>%
  ggplot(aes(SE.SEC.NENR.FE)) +
  geom_histogram()

joined %>%
  ggplot(aes(SE.SEC.NENR.MA)) +
  geom_histogram()

joined %>%
  ggplot(aes(SE.SEC.NENR.MA,SE.SEC.NENR.FE)) +
  geom_point()

joined %>%
  arrange(desc(SP.POP.TOTL)) %>%
  ggplot(aes(NY.GDP.PCAP.CD,secondary_enrollment)) +
  geom_point() +
  scale_x_log10() + 
  geom_text(aes(label=country), vjust=1, hjust=1, check_overlap=TRUE) +
  labs(x = "GDP per Capita",
       y = "Secondary School Enrollment")
```

```{r}
joined %>%
  arrange(desc(SP.POP.TOTL)) %>%
  ggplot(aes(student_ratio,secondary_enrollment)) +
  geom_point() +
  geom_text(aes(label=country), vjust=1, hjust=1, check_overlap=TRUE) +
  labs(x = "Student to Teacher Ratio",
       y = "Secondary School Enrollment")
```

#Examining the confounding variable
```{r}
lm(secondary_enrollment ~ student_ratio, joined) %>%
  summary()

lm(secondary_enrollment ~ student_ratio + log(NY.GDP.PCAP.CD), joined) %>%
  summary()

lm(secondary_enrollment ~ student_ratio + log2(NY.GDP.PCAP.CD), joined) %>%
  summary()


lm(secondary_enrollment ~ student_ratio + log10(NY.GDP.PCAP.CD), joined) %>%
  summary()

lm(secondary_enrollment ~ student_ratio + log2(NY.GDP.PCAP.CD) + region, joined) %>%
  aov() %>%
  summary()

joined %>%
  filter(!is.na(SE.XPD.TOTL.GD.ZS)) %>%
  lm(secondary_enrollment ~ student_ratio + log2(NY.GDP.PCAP.CD) + SE.XPD.TOTL.GD.ZS,
     data = .) %>%
  summary()

ggplot(joined, aes(SE.XPD.TOTL.GD.ZS, secondary_enrollment)) +
  geom_point()

joined %>%
    transmute(student_ratio,
              secondary_enrollment,
              log2_gdp = log2(NY.GDP.PCAP.CD)) %>%
  cor(use="pairwise.complete.obs")
```

#Appendix: Primary vs. Secondary Education
```{r}
secondary_primary_education %>%
  select(indicator,country,student_ratio) %>%
  mutate(indicator=snakecase::to_snake_case(indicator)) %>%
  spread(indicator,student_ratio) %>%
  mutate(ratio=secondary_education/primary_education) %>%
  ggplot(aes(primary_education, secondary_education)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() + 
  geom_text(aes(label=country), vjust=1, hjust=1, check_overlap=TRUE)
```